Structural replication of nanoporous media using procedural noise

### 2018-05-29 ms.tex
Added improved 1d noise figure, wrote more about model

\begin{figure}
\includegraphics[width=.5\textwidth]{1d_perlin.png}
\caption{1D simplex noise with 1 (blue), 3 (green) and 30 (red) octaves. }
\label{fig:1dperlin}
\end{figure}

\section{Model}
In this paper, we are interested in developing a model using simplex noise to reproduce
nanoporous structures that are statistically similar to physical simulations of porous media.
For a given data set and a model, we wish to estimate the parameters that best represent the data. 
This is performed through a full likelihood analysis in order to obtain the posterior
probability distribution of the parameters. 
The model measure of the analysis is the radial distribution function $g(r)$, for
which the parameters will be adjusted to maximise.
In addition, we also investigate the porosity and surface area of the system. 
The simulated system is confined to a cube of size $\SI{159} {\angstrom}^3$, but since
the procedural noise function is continuous on all scales, the box could be of arbitrary size.

\begin{figure*}
\includegraphics[width=.95\textwidth]{model_examples.png}
\caption{Examples of 3D simplex noise structures in nanoporous media. Left image:
low persistence (large scale dominates), middle image: medium persistence (small
scales starts dominating) and right image: high threshold (more atoms are discarded). }
\label{fig:model_example}
\end{figure*}
We construct a generic sum over procedural noise octaves with the expression: 
\begin{equation}
  M_1(\vec r |\bar \theta) = M_1(\vec r | \psi, \Omega, \nu) = \sum_{\omega=0}^{\Omega} \psi^{\omega -1}   N\big((\vec r + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
\label{eq:noisemodel1}
\end{equation}
where $\Omega$ is the number of octaves, $\psi$ the persistence (scale falloff),
$N$ a generic $N$-dimensional noise function, $\vec \sigma$ a predefined random
vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing
the initial largest scales in the model. For simplicity, $\bar \theta$ here denotes
the set of model parameters. For each mode, the frequency is thus multiplied, while
the overall amplitude is damped by a given persistence.

We then get, for a given position $\vec r$, a scalar value from the noise function.
Assume now that we have a system in molecular dynamics where all atomic positions are given.
A porous material is then generated by introducing a threshold $\tau$ which is used to decide whether or not an atom should be removed from the system.
A boolean function $\text{Delete}(\vec r)$ is defined as

\begin{align}
  \text{Delete}(\vec r) = 
  \begin{cases} 
    \text{True} & \text{if} \quad M_1(\vec r |\bar \theta) < \tau \\
    \text{False} & \text{if} \quad M_1(\vec r |\bar \theta) \geq \tau,
  \end{cases}
\end{align}
where we simply evaluate this function for every atom in the system.
The threshold $\tau$ is now added as a parameter to the model.
Three examples of this model are depicted in figure \ref{fig:model_example}, where we have
used $\Omega=6$ octaves and scale $\nu=0.01$ with threshold $\tau=0$. The left
image depicts low persistence ($\psi = 0.0$), where only the first octave dominates
(corresponding to large scales). The middle picture shows the same parameters with
a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
In the rightmost picture, the effect of the threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.

@@

The method used for calculating porosity of a porous media is the same as in \cite{gelb1998characterization}.
Surface area is obtained by creating a voxelized grid of the system with values being the distance from each voxel center to the nearest atom.
This is now a scalar field and we can apply marching cubes to triangulate an isosurface of the distance field.
This gives a trianglated representation of the porous media defined by a set of atomic positions.
As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in
atomistic simulations, because a clear definition of pore space is difficult.
We have chosen to define the pore space as the set of points being more than \SI{2}{\angstrom} to any atom.
This is large enough to contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such. 
The important note here is that the same definition of both surface and porosity is being used in all systems that are compared, physical simulations, procedural methods or data.

>>>>

\begin{figure}
\includegraphics[width=.5\textwidth]{1d_noise.pdf}
\caption{1-dimensional simplex noise with 1 (yellow), 3 (red) and 30 (blue) octaves. Adding more octaves adds high frequency features, but it still follows the 1 octave base value.}
\label{fig:1dperlin}
\end{figure}

\section{Model}
In this paper, we are interested in developing a model using simplex noise to reproduce
nanoporous structures that are statistically similar to physical simulations generated by molecular dynamics.
A series of statistical similar nanoporous materials are generated using molecular dynamics from which the model will be fitted to.
For a given data set and a model, we wish to estimate the parameters that best represent the data. 
This is performed through a full likelihood analysis in order to obtain the posterior probability distribution of the parameters. 
The model measure of the analysis is the radial distribution function (or pair correlation function) $g(r)$, a quantity that measures how the density varies as a function of the distance from an atom.
For low $r$, there is typically a peak at the equilibrium distance between nearest neighbor atoms in crystals or molecules, and several more peaks are often well defined as next nearest neighbors usually obey some symmetries as well.
The fitting procedure consists of measuring $g(r)$ for both the data and the model, and deviations contributes to the error $\chi^2$ (defined below).
In addition, we also investigate the porosity and surface area of the system after fitting since these are important physical quantities in such systems.

\begin{figure*}
\includegraphics[width=.95\textwidth]{model_examples.png}
\caption{Examples of 3D simplex noise structures in nanoporous media. Left image:
low persistence (large scale dominates), middle image: medium persistence (small
scales starts dominating) and right image: high threshold (more atoms are discarded). }
\label{fig:model_example}
\end{figure*}

Given a set of parameters $\bar \theta$, we construct a generic sum over procedural noise octaves with the expression:
\begin{align}
  M_1(\vec r |\bar \theta) &= M_1(\vec r | \psi, \Omega, \nu)\nonumber\\
  &= \sum_{\omega=1}^{\Omega} \psi^{\omega-1}   N\big((\vec r + \vec\sigma \cdot \omega)\nu \cdot 2^{\omega-1} \big),
\label{eq:noisemodel1}
\end{align}
where $\Omega$ is the number of octaves, $\psi$ the persistence (the amplitude at different scales),
$N$ a generic $N$-dimensional noise function (in this paper Perlin noise or simplex noise), $\vec \sigma$ a predefined random
vector that removes anisotropic artefacts and $\nu$ the frequency modifier representing
the initial largest scales in the model. For each octave, the frequency is doubled, while the overall amplitude is damped by a certain persistence $\psi$.

We then get, for a given position in physical space $\vec r$, a scalar value from the noise function.
Assume now that we have a bulk system in molecular dynamics where all atomic positions are given.
A porous material is then generated by introducing a threshold $\tau$ which is used to decide whether or not an atom should be removed from the system.
A boolean function $\text{Delete}(\vec r)$ is defined as

\begin{align}
  \text{Delete}(\vec r) = 
  \begin{cases} 
    \text{True} & \text{if} \quad M_1(\vec r |\bar \theta) < \tau \\
    \text{False} & \text{if} \quad M_1(\vec r |\bar \theta) \geq \tau,
  \end{cases}
\end{align}
where we simply evaluate this function for every atom in the system.
The threshold $\tau$ is now added as a parameter to the model.

Three examples of this model are depicted in figure \ref{fig:model_example}, where we have
used $\Omega=6$ octaves and scale $\nu=0.01$ with threshold $\tau=0$. The left
image depicts low persistence ($\psi = 0.0$), where only the first octave dominates
(corresponding to large scales). The middle picture shows the same parameters with
a higher persistence $(\psi = 0.5)$, where smaller scales become more dominating.
In the rightmost picture, the effect of the threshold $\tau = -0.3$ is shown to emphasize threshold cut-off.
The bulk system is a cubic system of amorphous SiO$_2$ with size \SI{159}{\angstrom}.

@@

The method used for calculating porosity of a porous media is the same as in \cite{gelb1998characterization}.
Surface area is obtained by creating a voxelized grid of the system with values being the distance from each voxel center to the nearest atom.
This is now a scalar field and we can apply marching cubes\cite{lorensen1987marching} to triangulate an isosurface of the distance field.
This gives a trianglated representation of the porous media defined by a set of atomic positions.
As discussed in \cite{gelb1998characterization}, the definition of surface area is ill-defined in
atomistic simulations, because a clear definition of pore space is difficult.
We have chosen to define the pore space as the set of points being more than \SI{2}{\angstrom} to any atom.
This is large enough to contribute to a dense pore matrix, but small enough to have most of the available pore space defined as such. 
The important note here is that the same definition of both surface and porosity is being used in all systems that are compared, physical simulations, procedural methods or data.
